---
title: "Residential Land in the ACT: How much is there?"
format:
  html:
    code-fold: true
---

```{r}
#| message: false
library(tidyverse)
library(sf)
sf_use_s2(FALSE)
```

The ACT government publishes a file which includes all blocks in the territory. The (somewhat limited) documentation for it is [here](https://www.data.act.gov.au/dataset/ACTGOV-BLOCK/h78d-hg25).

```{r}
#| cache: true
#| message: false
blocks_raw <- st_read('ACTGOV_BLOCK.geojson',quiet=TRUE)
blocks <- blocks_raw |>
  janitor::clean_names() |>
  filter(current_lifecycle_stage %in% c('APPROVED','REGISTERED','OCCUPIED')) |>
  mutate(area = st_area(geometry)) 
```

With this data in hand, we can sum up all land contained in blocks contained in zones RZ1-RZ5:

```{r}
#| message: false
#| warning: false
rz_blocks <- blocks |>
  filter(str_detect(land_use_policy_zones,'^RZ[1-5]:')) |> filter(!str_detect(land_use_policy_zones,';'))  #this filters out all blocks listed as being in more than one zone (these are not large areas, I'm not sure exactly what they are)
total_areas <- rz_blocks |> 
  summarise(total_area = sum(area)) |>
  mutate(total_area = units::set_units(total_area,'hectare'))
total_areas |> as_tibble()
```

I have excluded blocks with more than one zone listed (I'm not sure what's going on in those cases).

To fully answer the question of how much residential land there is in the ACT, we'd have to account for CZ zones (although land used for residential purposes on CZ* land would be trivial compared to residential RZ land), and define exactly what is meant by "residential land".